---
title: "GOA pollock"
# format: html
editor: visual
format:
  html: 
    page-layout: full
    fig-format: svg
    fig-dpi: 400
    fig-width: 80
    fig-height: 50
---

## Background

*The Groundfish Assessment Program is migrating to new database tables (housed in the `GAP_PRODUCTS` schema) starting in the 2024 groundfish stock assessment cycle. Here we compare GOA bottom trawl survey (BTS) biomass, size frequencies, and age comps for the GOA pollock assessment.*

## Setup

```{r}
#| echo: false
#| output: false

libs <- c("tidyverse", "odbc", "keyring", "dbplyr", "kableExtra", "ggridges", "janitor")
if(length(libs[which(libs %in% rownames(installed.packages()) == FALSE )]) > 0) {
  install.packages(libs[which(libs %in% rownames(installed.packages()) == FALSE)])}
lapply(libs, library, character.only = TRUE)

theme_set(theme_bw(base_size = 15) + 
            theme(panel.border = element_blank(), 
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(), 
                  axis.line = element_line(colour = "black")))

akfin <- DBI::dbConnect(odbc::odbc(), "akfin", uid = keyring::key_list("akfin")$username, 
                        pwd = keyring::key_get("akfin", keyring::key_list("akfin")$username))

```

## *Species codes*

*This is the RACE `species_code`* for walleye pollock.

```{r}
#| output: false
my_spp_codes <- 21740
```

## *New area IDs*

\*Instead of hosting various survey data in different database tables, the new `GAP_PRODUCTS` uses `SURVEY_DEFINITION_ID` and `TYPE` to identify the survey and the structure or aggregation of data desired by the user. For example, in the Atka mackerel assessment, we pull biomass data at the regulatory area level.

[Documentation for GAP Bottom Trawl Surveys and Survey Definition IDs] (https://afsc-gap-products.github.io/gap_products/content/intro-survey-background.html)

```{r}
lkup <- dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_area')) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::filter(survey_definition_id %in% c(47)) %>%
  dplyr::collect()
```

```{r}
#| echo: false
lkup %>% 
  dplyr::arrange(area_type) %>% 
  kableExtra::kable("html") %>% 
    kableExtra::kable_styling() %>% 
    kableExtra::scroll_box(width = "100%", height = "200px")
```

## Biomass

### *Old data from Answers:*

Original data: [AKFIN Answers Biomass by NMFS Reporting Area - GOA](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard)

This table contains area swept estimates of biomass and population numbers by NMFS or INPFC Fisheries Reporting Areas for the Gulf of Alaska (GOA) Biennial Bottom Trawl Surveys.

The trawl survey regions generally correspond to the NMFS or INPFC Fisheries Reporting Areas. GOA stratum estimates are summed into the NMFS or INPFC Fisheries Reporting Area and include Shumagin, Chirikof, Kodiak, Yakutat, and Southeastern areas for the GOA Survey.

Note that the boundary for the NMFS Fisheries Reporting Area differs from the boundary for the INPFC reporting area (see splitfractionsgoa).

```{r}
#| output: false
old <- read_csv('data/answers_goapk_nmfsarea_biomass.csv') %>% 
  janitor::clean_names() %>%
  dplyr::select(area_name = nmfs_reporting_area, area_id = summary_area, 
                species_code, year, biomass_mt = area_biomass, 
                biomass_var = area_biomass_var)
```

```{r}
#| echo: false
old %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *New query:*

```{r}
area_ids <- lkup %>% filter(area_type == "INPFC") %>% pull(area_id)

new <- dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_biomass')) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(survey_definition_id, area_id, species_code, year, biomass_mt, biomass_var) %>% 
  dplyr::filter(species_code %in% my_spp_codes & area_id %in% area_ids & year >= 1990) %>% 
  dplyr::collect() %>% 
  dplyr::left_join(lkup %>% select(survey_definition_id, area_id, area_name))
```

```{r}
#| echo: false
new %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *Comparisons:*

```{r}
#| echo: false
compare <- new %>% 
  mutate(schema = "akfin.gap_products") %>% 
  select(schema, year, area_name, area_id, species_code, biomass_mt, biomass_var) %>% 
  bind_rows(old %>% 
              mutate(schema = "akfin.answers") %>% 
              select(schema, year, area_name, area_id, species_code, biomass_mt, biomass_var)) %>% 
  mutate(cv = ifelse(biomass_mt > 0, sqrt(biomass_var) / biomass_mt, NA))
```

GOA BTS Biomass (+/- 1 SD), red = old data, blue = new data:

```{r,fig.width=13,fig.height=11}
#| echo: false
ggplot(compare, aes(x = factor(year), y = biomass_mt, fill = schema)) +
  geom_bar(stat = 'identity', position = position_dodge2(width = 0.5, preserve = "single", padding = 0)) +
  geom_errorbar(aes(ymin = biomass_mt - sqrt(biomass_var), ymax = biomass_mt + sqrt(biomass_var)),
                position = position_dodge2(width = 0.5, preserve = "single", padding = 0)) +
  facet_wrap(~area_name, scale = 'free_y') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = NULL)
```

```{r}
#| echo: false
output <- compare %>% 
  pivot_wider(id_cols = c(year, area_name, species_code), names_from = schema, values_from = biomass_mt, values_fill = 0) %>% 
  mutate(percent_change = (akfin.gap_products-akfin.answers)/akfin.answers * 100,
         difference = akfin.gap_products-akfin.answers,
         variable = 'mean') %>% 
  bind_rows(compare %>% 
              pivot_wider(id_cols = c(year, area_name, species_code), names_from = schema, values_from = cv, values_fill = 0) %>% 
              mutate(percent_change = (akfin.gap_products-akfin.answers)/akfin.answers * 100,
                     difference = akfin.gap_products-akfin.answers,
                     variable = 'cv')) %>% 
  arrange(desc(variable), desc(abs(difference))) 

mean_diff <- output %>% filter(variable == 'mean' & difference == max(abs(difference)))
```

There were no substantive differences in biomass found. The largest difference in biomass between the two data sets was `r round(mean_diff$difference, 4)` t, which occurred in `r mean_diff$area_name` in `r mean_diff$year`. This was only a `r round(mean_diff$percent_change,4)`% change in reported biomass. These differences are likely attributable to rounding or truncation errors.

Comparison of biomass point estimates, largest differences shown first:

```{r}
#| echo: false
output %>%
  filter(variable == 'mean') %>% 
 kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

Comparison of biomass CVs, largest differences shown first:

```{r}
#| echo: false
output %>%
  filter(variable == 'cv') %>% 
 kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

## EGOA Fractions

### *Old data from Answers:*

Original data: [Biomass Fractions in Eastern GOA](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard&PortalPath=%2fshared%2fStock%20Assessment%2f_portal%2fStock%20Assessment&Page=RACE%20-%20Biomass%20Fractions&Done=Dashboard%26PortalPath%3d%252fshared%252fStock%2520Assessment%252f_portal%252fStock%2520Assessment%26Page%3dRACE%2520Survey%2520%26ViewState%3dgj9fftdbmt2o13d8auqsu1f9mm)

This report contains post-stratified biomass estimates for those strata bordering the 140° W longitude. Because the trawl survey regional boundary between the Southeastern and Yakutat in the Gulf of Alaska differs from the NMFS Fisheries Reporting Area, stations are reallocated to the latter and new biomass estimates are made for the Yakutat and Southeast regions.

```{r}
#| output: false
old <- read_csv('data/answers_goapk_egoa_fractions.csv') %>% 
  janitor::clean_names() %>%
  dplyr::select(year, management_group, total_biomass = total_biomass_mt, 
                east_biomass = eastern_biomass_mt, east_fraction = eastern_fraction, 
                west_biomass = western_biomass_mt, west_fraction = western_fraction)
```

```{r}
#| echo: false
old %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *New query:*

```{r}
new <- dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_split_fractions')) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(-akfin_load_date) %>% 
  dplyr::filter(management_group %in% c('walleye pollock')) %>% 
  dplyr::collect()
```

```{r}
#| echo: false
new %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *Comparisons:*

```{r}
#| echo: false
compare <- new %>% mutate(schema = "akfin.gap_products") %>% 
  bind_rows(old %>% mutate(schema = "akfin.answers")) %>% 
  tidyr::pivot_longer(cols = -c(management_group, schema, year))
```

[A github issue submitted 3/12/2024 about remaining 1980s EGOA fraction data and the fact that 2023 EGOA fractions haven't been uploaded yet](https://github.com/afsc-gap-products/gap_products/issues/32). Otherwise there were no differences in the eastern GOA fractions detected (differences were less then 1e-10).

```{r,fig.width=13,fig.height=11}
#| echo: false
ggplot(compare, aes(x = year, y = value, col = schema, lty = schema, 
                    shape = schema, group = schema)) +
  geom_point() +
  geom_line() +
  facet_wrap(~name, scale = 'free_y') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = NULL, x = NULL)

```

```{r}
#| echo: false
output <- compare %>% 
  pivot_wider(names_from = schema, values_from = value) %>% 
  mutate(percent_change = (akfin.gap_products-akfin.answers)/akfin.answers * 100,
         difference = akfin.gap_products-akfin.answers) %>% 
  arrange(desc(abs(difference))) 
```

```{r}
#| echo: false
output %>%
  kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

## Size frequencies

### *Old data from Answers:*

Original data: [AKFIN Answers Size Composition by Stratum - GOA](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard)

This report contains the size composition of stratum population estimates in terms of number of fish by size and by sex.

For each haul, the length and sex frequency proportions are calculated from the observations and then weighted by the cpue numbers. For any station with a positive cpue, but missing length and sex observations, the sex-length frequencies for the positive stratum stations observations are substituted for missing observations. Once the sex-length proportions for each station is calculated, the proportions are weighted by the station cpue to the total cpue and are then multiplied by population numbers estimate for the stratum to yield the abundance of males, females, and unknown sex at length for the stratum.

```{r}
#| output: false
old <- read_csv('data/answers_goapk_stratum_sizecomps.csv') %>%
  janitor::clean_names() %>% 
  dplyr::select(area_id = stratum, species_code, year, 
                length = length_mm, population_count = total_number, 
                area_name = stratum_description)
```

```{r}
#| echo: false
old %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *New query:*

```{r}
area_ids <- lkup %>% filter(area_type == "STRATUM") %>% pull(area_id)

new <- dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_sizecomp')) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(survey_definition_id, area_id, species_code, year, length = length_mm, population_count) %>% 
  dplyr::filter(survey_definition_id == 47 &
                  species_code %in% my_spp_codes & 
                  area_id %in% area_ids) %>% 
  dplyr::collect() %>%
  dplyr::left_join(lkup %>% 
                     filter(area_id %in% area_ids) %>%
                     distinct(survey_definition_id, area_id, area_name), 
                   by = join_by(survey_definition_id, area_id)) %>% 
  dplyr::select(-survey_definition_id)
```

```{r}
#| echo: false
new %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *Comparisons:*

```{r}
#| echo: false
# all areas combined
compare <- new %>% mutate(schema = "akfin.gap_products") %>% 
  bind_rows(old %>% mutate(schema = "akfin.answers")) %>% 
  group_by(schema, year, length) %>% 
  dplyr::summarise(population_count = sum(population_count))

# areas specific
# compare_a <- new %>% mutate(schema = "akfin.gap_products") %>% 
#   bind_rows(old %>% mutate(schema = "akfin.answers")) %>% 
#   group_by(schema, year, area_name, length) %>% 
#   dplyr::summarise(population_count = sum(population_count))
```

```{r,fig.width=14,fig.height=8}
#| echo: false
ggplot(compare, aes(x = length, y = factor(year), height = population_count, fill = schema)) +
  ggridges::geom_density_ridges(stat = "identity", scale = 4, alpha = 0.5) +
  theme(legend.position = "top") +
labs(y = NULL) 
```

```{r}
#| echo: false
output <- compare %>% 
  pivot_wider(id_cols = c(year, length), names_from = schema, values_from = population_count) %>% 
  mutate(percent_change = (akfin.gap_products-akfin.answers)/akfin.answers * 100,
         difference = akfin.gap_products-akfin.answers) %>% 
  arrange(desc(abs(difference))) 

max_diff <- output %>% filter(difference == max(abs(difference)))
```

[A github issue submitted 3/12/2024 about the 2025 size comp data in the GAP_PRODUCTS data](https://github.com/afsc-gap-products/gap_products/issues/33). There are some small increases in size frequencies in 1993 (<3%) that affects a few of the smaller size length bins. I don't know what is contributing to these increases. The

```{r}
#| echo: false
output %>%
 kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

## Specimen data 

### *Old data from email:*

From: Ned Laman - NOAA Federal <ned.laman@noaa.gov>
Date: Tue, Oct 31, 2023 at 8:14 AM
Subject: Re: One more small request

```{r}
#| output: false
old <- read_csv('data/176202101201.csv') %>%
  janitor::clean_names() %>% 
  mutate(year = lubridate::year(date_collected))
```

### *New query:*

You should ask GAP the best way to query specimen data to get your sample sizes. Here was my best attempt (not good enough).

```{r}
new <- dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_specimen')) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(hauljoin, specimen = specimen_id, species_code, length = length_mm, 
                weight = weight_g, age, sex, age_method = age_determination_method) %>% 
  dplyr::left_join(dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_haul')) %>%
                     dplyr::rename_all(tolower) %>% 
                     dplyr::select(cruisejoin, hauljoin, haul, date_collected = date_time_start, 
                                   latitude = latitude_dd_start, longitude = longitude_dd_start),
                   by = join_by(hauljoin)) %>% 
  dplyr::left_join(dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_cruise')) %>%
                     dplyr::rename_all(tolower) %>% 
                     dplyr::select(cruisejoin, year, vessel = vessel_id, survey_definition_id),
                   by = join_by(cruisejoin)) %>% 
  dplyr::filter(survey_definition_id == 47 & species_code %in% my_spp_codes) %>% 
  dplyr::collect() %>% 
  dplyr::select(-survey_definition_id)
```

```{r}
#| echo: false
new %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *Comparisons:*

The age method codes are different in the different tables. Not going to worry about that for now, but maybe you want to look into it later.
```{r}
#| echo: false
compare <- new %>% mutate(schema = "akfin.gap_products") %>% select(-age_method)%>% 
  bind_rows(old %>% mutate(schema = "email.gap") %>% select(-age_method)) 

nlens <- compare %>% 
  filter(!is.na(length)) %>% 
  group_by(schema, year) %>% 
  summarise(n = length(length)) %>% 
  pivot_wider(names_from = schema, values_from = n)

nages <- compare %>% 
  filter(!is.na(age)) %>% 
  group_by(schema, year) %>% 
  summarise(n = length(age)) %>% 
  pivot_wider(names_from = schema, values_from = n)

```

I'm clearly missing a filter because the 2021 sample sizes are not matching up. 

Sample sizes for survey lengths
```{r}
#| echo: false
nlens %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

Sample sizes for survey ages
```{r}
#| echo: false
nages %>% kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

## Age comps

### *Old data from email:*

From: Ned Laman - NOAA Federal <ned.laman@noaa.gov>
Date: Tue, Oct 31, 2023 at 8:14 AM
Subject: Re: One more small request

```{r}
#| output: false

old <- read_csv('data/aigoa_work_data.goa2021pollockagelengthbyarea.csv') %>%
  janitor::clean_names() %>% 
  mutate(year = 2021)

old <- old %>% 
  mutate(age=pmin(age,10)) %>%
  filter(age %in% 1:10, sex==9, summary_area %in% c(919,929,939)) %>%
  group_by(year, age) %>% summarize(num=sum(agepop), .groups='drop') %>%
  mutate(prop=num/sum(num),
         version='email.gap')
```
### *New query:*

```{r}
# original INPFC areas used for age comps
area_ids <- lkup %>% filter(area_id %in% c(919,929,939)) %>% pull(area_id)
#   area_type area_id description               
# 1 INPFC         919 Shumagin Area 1-500 meters
# 2 INPFC         929 Chirikof Area 1-500 meters
# 3 INPFC         939 Kodiak Area 1-500 meters 

# looks like only AREA_TYPE as STRATUM or REGION are available. not sure why. here's my attempt at recreating the original area_ids
area_ids <- lkup %>% filter(area_type == 'STRATUM' & 
                              !grepl('Southeastern|Yakutat|Southeast',description) & 
                              depth_max_m <= 500) %>% pull(area_id)

new <- dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_agecomp')) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(year, survey_definition_id, species_code, summary_area = area_id, 
                sex, age, agepop = population_count, 
                meanlen = length_mm_mean, sdev = length_mm_sd) %>% 
  dplyr::filter(survey_definition_id == 47 & species_code %in% my_spp_codes &
                  summary_area %in% area_ids) %>%
  dplyr::collect() %>% 
  dplyr::select(-survey_definition_id, -species_code)

```

```{r}
#| echo: false
comps <- new %>% 
  mutate(age=pmin(age,10)) %>%
  filter(age %in% 1:10, sex==3) %>%
  group_by(year, age) %>% 
  summarize(num=sum(agepop), .groups='drop') %>%
  group_by(year) %>% 
  mutate(prop=num/sum(num),
         version='akfin.gap_products') 

compare <- comps %>% filter(year == 2021) %>% bind_rows(old)

comps %>% 
  kableExtra::kable("html") %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "200px")
```

### *Comparisons:*

Haha I'm so glad your age comps don't look like this... I really don't know why this didn't work. Sry! 
```{r,fig.width=8,fig.height=8}
#| echo: false
ggplot(compare, aes(x = age, y = factor(year), height = prop, fill = version)) +
  ggridges::geom_density_ridges(stat = "identity", scale = 4, alpha = 0.5) +
  theme(legend.position = "top") +
labs(y = NULL) 
```
## Summary of results

Biomass, size frequencies, and EGOA fractions look good.

Near-term requests/questions for GAP:
1) a query to reproduce your 2021 age comps
2) an equivalent GAP_PRODUCT query that reproduces your 2021 age comps
3) why only AREA_TYPE as STRATUM or REGION are available in the AKFIN_AGECOMP table
4) an example query for the SPECIMEN table that includes joins to the CRUISE AND HAUL tables that you can use for your sample size tables

Long-term requests for GAP:
Currently the portion of the EGOA west of 140W are not represented in the age comps. Is there a way to include those age data?